---
title: "Programming Fundamentals"
author: "Marina Wyss"
date: "8/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Agenda
* Data types
    * Numeric/integers
    * Strings
    * Booleans
    * Factors
    * Vectors
    * Matricies
    * Data frames
    * Lists
* Comparison Operators
* Logical Operators
* if, else if, else statements
* for loops
* while loops
* functions
* Exercises
* Solutions

<i> Some examples in this worksheet are inspired by Pierian Data's Python and R courses on Udemy.</i><br>
<br>
This worksheet is available in both <b>R</b> and <b>Python</b> (R below of course), with some minor differences. We think it's important to become familiar with Python early on, since it is also a very popular language and will be necessary to know if you want to pursue a career with data. Some key points about the two:
<ul>
<li> R comes from stats, Python comes from computer science</li>
<li> So, R has many statistical methods that aren't available or are more challenging in Python, and Python has some machine learning/AI capabilities that aren't possible in R.</li>
<li> Python is a full-on programming language, with options far beyond data analysis.</li>
<li> R is more common in academia.</li>
</ul>
Neither is "harder" to learn than the other - it's more a matter of what you're used to. It makes sense to practice both as much as possible.

### numeric/integers
Numerics: can be whole numbers or doubles/floats with decimals
Integers: only whole numbers, though often R just considers them numerics
Often the distinction isn't really important.
```{r}
# numeric
four <- 2 + 2
class(four)

flo <- 2.5 - 1.7
class(flo)

# include L at the end to tell R it's an integer and not normal numeric
fourInt <- 2L + 2L
class(fourInt)
```

### strings
Can use single or double quotes.
Can be characters or numeric, but will be stored as characters.
```{r}
hi <- "Hello world"
class(hi)

numString <- "12345"
class(numString)
```

### booleans
True/false. Must be in caps. Can be either the whole word or just T/F, though the whole word is better form.
```{r}
TRUE
T

FALSE
F
```

### factors
Categorical variables.
```{r}
# variable gender with 10 "male" entries and 10 "female" entries 
gender <- c(rep("male", 10), rep("female", 10)) 
gender <- factor(gender) 

# or as one line
gender <- as.factor(c(rep("male", 10), rep("female", 10)))

str(gender)

# can also create an ordered factor for ordinal variables
sizes <- c(rep("large", 10), rep("medium", 10), rep("small", 10)) 
sizes <- ordered(sizes)

str(sizes)
```

### vectors
One-dimensional vector that must have all the same element type
```{r}
vec1 <- 10:50
vec1

vec2 <- rep("meow", 10)
vec2
```

### matrices
A matrix is a collection of data elements arranged in a two-dimensional rectangular layout. All columns in a matrix must have the same class (numeric, character, etc.) and the same length.
```{r}
myMat <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9), 
                 nrow = 3, 
                 ncol = 3, 
                 byrow = TRUE)
myMat
```

### data frames
Basically a "spreadsheet". Two-dimensional. Key is that they can have different data types. 
```{r}
names <- c('Tom', 'Nick', 'Sally', 'Joe')
ages <- c(20, 21, 19, 18)
myData <- data.frame(ages, names)
names(myData) <- c("Age", "Name")

myData
```

### lists
An ordered collection of objects (components). A list allows you to gather a variety of (possibly unrelated) objects under one name. For example, could have a few data frames, a matrix, etc.
```{r}
myList <- list(myData, myMat, vec2, sizes)
myList
```

### Comparison Operators
```{r}
1 > 2
1 < 2
1 >= 1
1 <= 4
1 == 1
'hi' == 'bye'
```

### Logical Operators
```{r}
# all statements must be true to return TRUE when using "and"
(1 > 2) & (2 < 3)

# only one statement must be true to return TRUE when using "or", denoted with |
(1 > 2) | (2 < 3)

(1 == 2) | (2 == 3) | (4 == 2)
```

### if, else if, else statements
```{r}
if (1 < 2){
  print('yep!')
}

if (1 < 2){
    print('first')
} else {
    print('last')
}

if (1 == 2){
  print('first')
} else if(3 == 3) {
  print('middle')
} else {
  print('last')
}
```

### for loops
for each item, perform some action.
```{r}
nums <- 1:5

for (item in nums){
  print(item)
}

for (item in nums){
  print("Yep")
}

# can use whatever word to iterate over
for (cake in nums){
  print(cake + cake)
}
```

### while loops
while loops are a way to have your program continuously run some block of code until a condition is met (or in other words, while a certain condition is true). 

A major concern when working with a while loop is to make sure that at some point the condition should become true, otherwise the while loop will go forever! 
```{r}
i <- 1

while (i < 5){
  print(paste0("i is: ", i))
  i <- i + 1
}
```

### functions
Formally, a function is a useful device that groups together a set of statements so they can be run more than once. They can also let us specify parameters that can serve as inputs to the functions.

On a more fundamental level, functions allow us to not have to repeatedly write the same code again and again. 
```{r}
# basic structure
functionName <- function(someInput){
  # some action to perform
}

# example 1
square <- function(x){
  print(x^2)
}

square(2)

# example 2
greeting <- function(name){
    print(paste('Hi ', name))
}

greeting("dude")

# example 3
addNum <- function(num1, num2){
    print(num1 + num2)
}
addNum(5, 10)
```
Now, here's a more complicated example. It will return the lesser of two given numbers if both numbers are even, but returns the greater if one or both numbers are odd. <br>
For example: 
* lesser_of_two_evens(2, 4) would return 2
* lesser_of_two_evens(2, 5) would return 5
```{r}
lesser_of_two_evens <- function(a, b){
  if ((a %% 2) == 0 & (b %% 2) == 0){
    print(min(a, b))
  } else {
    print(max(a, b))
  }
}

lesser_of_two_evens(2, 4)
lesser_of_two_evens(2, 5)
```

